Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-03-24 demystify.tex
updates

\section{How much time did you spend looking at my application?}\label{sec:time-spent}

Around 15 minutes on average (less than 10 mins for definite rejection; 20+ mins for an acceptance). Note I spent more time in my earlier years, but now I have more experience and can quickly tell if an application is good or bad.

The evaluation system allows us to download a single pdf that consists of a summary page (e.g., schools, GPA, standard test scores), transcripts, test results, LoRs, CV, SOP, and writing samples (if any). I typically review the application in this order.  I also make notes on what I've seen, using them to support my evaluation.

@@

At this point I have enough to make up my decision, which I enter to the evaluation system. We have several options for decision, but typically I just use \emph{reject} and \emph{admit with full funding}.  For the latter, I also mention funding sources (e.g., RA from prof. X or TA). I can also recommend very strong students for the University Presidential fellowship.  Finally, the system has text box where I put my comments to support my decision.

After I submit my evaluation, the adcom chair will look all evaluations for that application and make a decision (almost always following our suggestion).  If there are large discrepancies such as one vote reject but another full funding acceptance (such cases are rare, but do happen), the chair might ask us to coordinate and re-evaluate the application if necessary.

@@

\section{Member of an underrepresented or LGBTQ+ groups}\label{sec:urm}
A common question is whether you should mention that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} or LGBTQ+ group (, e.g., in your SOP).
This is a strong concern for international (and even domestic) students, who may have encountered discrimination in their countries or personal experiences.
%This question arises often for international (and even domestic) students. The main worry is that they might get discriminated against, as often happen in their countries or personal experiences. %, and (ii) mentioning this might make them look like they are asking for special treatment.

@@

Ultimately, this is a personal choice and you do what you feel comfortable. Seeking guidance from your close professors and collaborators is a good idea. However, it is important to recognize that diversity is strongly valued and celebrated within CS programs in the US, and that you should feel empowered to share your unique experiences and perspectives.

>>>>

\section{How much time do you spend looking at an application?}\label{sec:time-spent}

Around 15 minutes on average (less than 10 mins for definite rejection; 20+ mins for acceptance). In earlier years I spent more time, but now I can quickly tell if an application is good or bad.

The evaluation system allows for downloading a single pdf that consists of a summary page (e.g., schools, GPA, standard test scores), transcripts, test results, LoRs, CV, SOP, and writing samples (if any). I typically review the application in this order.  I also make notes on what I've seen, using them to support my evaluation.

@@

At this point I have enough to make up my decision, which I enter to the evaluation system. Typically my decisions are either \emph{reject} or \emph{admit with full funding}.  For the latter, I also mention funding sources (e.g., RA from prof. X or TA). I also include my comments to support the decision. Finally, I also recommend very strong students for the University Presidential fellowship (\autoref{sec:fellowships}).

The adcom chair will look all evaluations for each application and makes a decision (likely following the decision of the reviewers).  If there are discrepancies such as one voted reject but another full funding acceptance (such cases are rare, but do happen), the chair might ask the reviewers to coordinate to reach a concensus.

@@

\section{Member of an underrepresented or LGBTQ+ groups}\label{sec:urm}
A common question is whether you should mention that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} or LGBTQ+ group (e.g., in your SOP).
This is a strong concern for international (and even domestic) students, who may have encountered discrimination in their countries or personal experiences.
%This question arises often for international (and even domestic) students. The main worry is that they might get discriminated against, as often happen in their countries or personal experiences. %, and (ii) mentioning this might make them look like they are asking for special treatment.

@@

Ultimately, this is a personal choice and you do what you feel comfortable.
If you do not feel comfortable disclosing this, you should not feel pressured to do so and you should also ask your LoR writers to not disclose this information.
Note that you can still talk about your experiences and perspectives without mentioning your identity.
If uncertain, you can seek additional guidance from your close professors and collaborators However, it is important to recognize that diversity is strongly valued and celebrated within CS programs in the US, and that you should feel empowered to share your unique experiences and perspectives.
